Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-10-01 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in exp setup

## Overview

@@

```{r results='asis', eval = FALSE}
data.frame(j = c(2, 3, 6, 18),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               align = "rl",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $j$ and $X_{raw}$', 
               label = "parameter-table" ) %>%
  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

The values of $j$ was chosen so that different shapes of non-linearity were included in the residual plot. As shown in Figure \ref{fig:different-shape-of-herimite}, these include "U" shape, "S" shape, "M" shape and "Triple-U" shape.

@@

<!-- The range of $\sigma$, which is a factor controlling the strength of the signal, was chosen so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous. The example residual plots are given in Figure \ref{fig:different-sigma}. -->

Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density. And the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-lineup} shows one of the lineups used in experiment I. This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

```{r example-lineup, fig.cap="Lineup 24 in experiment I. Can you spot the most different plot? \\label{fig:example-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("Left-triangle shape")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Butterfly shape")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("Right-triangle shape")) +
  theme_light() -> p3

patchwork::wrap_plots(p1, p2, p3, ncol = 3)
```

@@

### Treatments

To control the strength of the signal, two parameters were used including $\sigma$ and $b$ for controlling the standard deviation of the error term in experiment I and II respectively.

<!-- > Experiment I -->

<!-- -   Treatments for nonlinearity experiment (plots summarising levels) -->

<!-- It was generated for $n = 1$, $\sigma = 1$, $j = 1$, $X_{raw} \sim N$. For this lineup, x out of k observers picked the actual data plot. Each lineup is evaluated by 5 different subjects to provide reasonable estimates of the p-value. -->

>>>>

## Experiment overview

@@

```{r results='asis'}
data.frame(j = c(2, 3, 6, 18),
           x_dist = c("$U(-1, 1)$", "$N(0, 0.3^2)$", "$lognormal(0, 0.6^2)/3$", "$U\\{1, 5\\}$")) %>%
  knitr::kable(col.names = NULL,
               align = "rl",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Parameter values for $j$ and $X_{raw}$', 
               label = "model-parameter-table" ) %>%
  kableExtra::add_header_above(c("Order of Hermite polynomial\n(sym1)", "Distribution of sym4\n")) %>%
  as.character() %>%
  gsub("sym1", "$j$", .) %>%
  gsub("sym4", "$X_{raw}$", .) %>%
  cat()
  # gsub("\\\\\\$", "$", .) %>%
  # cat()
```

The values of $j$ was chosen so that distinct shapes of non-linearity were included in the residual plot. A greater value of $j$ will result in a curve with more turning points. As shown in Figure \ref{fig:different-shape-of-herimite}, it includes "U" shape, "S" shape, "M" shape and "Triple-U" shape. It is expected that the "U" shape will be the easiest one to detect because complex shape tends to be concealed by cluster of data points.

@@

Four different distribution were used to generate $X_{raw}$ as shown in Figure \ref{fig:different-dist}. The uniform and the normal distribution are symmetric and commonly assumed in statistical models. The adjusted log-normal distribution provides skewed density, while the discrete uniform distribution provides discreteness in residual plot, which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-poly-lineup} shows one of the lineups used in experiment I. This lineup was produced under $j = 6$ and $X_{raw} \sim N(0.0.3^2)$. The actual data plot location was four. All five subjects correctly identified the actual data plot for this lineup.

```{r example-poly-lineup, fig.cap="Lineup poly-24 in experiment I. Can you spot the most different plot? \\label{fig:example-poly-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["poly_24"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

@@

VI_MODEL$plot(ori_dat) +
  ggtitle(quote("a = -1: Left-triangle shape")) +
  theme_light() -> p1

VI_MODEL$plot(heter_model(a = 0, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("a = 0: Butterfly shape")) +
  theme_light() -> p2

VI_MODEL$plot(heter_model(a = 1, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, b = 128)$gen(300, computed = select(ori_dat, x, e))) +
  ggtitle(quote("a = 1: Right-triangle shape")) +
  theme_light() -> p3

patchwork::wrap_plots(p1, p2, p3, ncol = 2)
```

```{r example-heter-lineup, fig.cap="Lineup heter-169 in experiment II. Can you spot the most different plot? \\label{fig:example-heter-lineup}"}
VI_MODEL$plot_lineup(vi_lineup[["heter_169"]]$data, 
                     remove_grid_line = TRUE, 
                     theme = theme_light(),
                     remove_axis = TRUE)
```

An example lineup of this model is shown in Figure \ref{fig:example-heter-lineup} with $a = -1$ and $X_{raw} \sim U(-1, 1)$. The actual data plot location was 15. 8 out of 11 subjects correctly identified the actual data plot for this lineup.

@@

### Controlling the strength of the signal

Except for the parameters summarised in Table \ref{tab:model-parameter-table}, three additional parameters $n$, $\sigma$ and $b$ were used to control the strength of the signal so that different difficulty levels of lineups were generated, and therefore, the estimated power curve would be smooth and continuous.

In experiment I, II and III, three different sample sizes were selected (n = 50, 100, 300). In general, with fewer data points drawn in a residual plot, the chance of detecting the visual pattern will decrease. This can be observed from Figure \ref{fig:different-n}.

```{r different-n, fig.cap="Three different values of $n$ are used in experiment I, II and III to control the strength of the signal."}
ori_dat <- poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(300)

VI_MODEL$plot(ori_dat) + 
  ggtitle(quote(n~"="~300)) + 
  theme_light() -> p1

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:50,], x, e))) +
  ggtitle(quote(n~"="~50)) + 
  theme_light() -> p2

VI_MODEL$plot(poly_model(shape = 2, x = {raw_x <- rand_uniform(-1, 1); closed_form(~stand_dist(raw_x))}, sigma = 1.5)$gen(50, computed = select(ori_dat[1:100,], x, e))) +
  ggtitle(quote(n~"="~100)) + 
  theme_light() -> p3

patchwork::wrap_plots(p2, p3, p1, ncol = 2)
```

Four different levels of $sigma$

### 2022-10-01 paper_comparison/paper_comparison/paper_comparison.tex
fill in exp setup

\hypertarget{overview}{%
\subsection{Overview}\label{overview}}

@@

The values of \(j\) was chosen so that different shapes of non-linearity
were included in the residual plot. As shown in Figure
\ref{fig:different-shape-of-herimite}, these include ``U'' shape, ``S''
shape, ``M'' shape and ``Triple-U'' shape.

@@

Four different distribution were used to generate \(X_{raw}\) as shown
in Figure \ref{fig:different-dist}. The uniform and the normal
distribution are symmetric and commonly assumed in statistical models.
The adjusted log-normal distribution provides skewed density. And the
discrete uniform distribution provides discreteness in residual plot,
which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-lineup} shows one of the lineups used in
experiment I. This lineup was produced under \(j = 6\) and
\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
five subjects correctly identified the actual data plot for this lineup.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-lineup-1} \caption{Lineup 24 in experiment I. Can you spot the most different plot? \label{fig:example-lineup}}\label{fig:example-lineup}
\end{figure}

@@

@@

\hypertarget{treatments}{%
\subsubsection{Treatments}\label{treatments}}

To control the strength of the signal, two parameters were used
including \(\sigma\) and \(b\) for controlling the standard deviation of
the error term in experiment I and II respectively.

>>>>

\hypertarget{experiment-overview}{%
\subsection{Experiment overview}\label{experiment-overview}}

@@

\caption{\label{tab:model-parameter-table}Parameter values for $j$ and $X_{raw}$}
\centering
\begin{tabular}[t]{rl}
\toprule
\multicolumn{1}{c}{\makecell[c]{Order of Hermite polynomial\\($j$)}} & \multicolumn{1}{c}{\makecell[c]{Distribution of $X_{raw}$\\}} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2}
2 & $U(-1, 1)$\\
3 & $N(0, 0.3^2)$\\
6 & $lognormal(0, 0.6^2)/3$\\
18 & $U\{1, 5\}$\\
\bottomrule
\end{tabular}
\end{table}

The values of \(j\) was chosen so that distinct shapes of non-linearity
were included in the residual plot. A greater value of \(j\) will result
in a curve with more turning points. As shown in Figure
\ref{fig:different-shape-of-herimite}, it includes ``U'' shape, ``S''
shape, ``M'' shape and ``Triple-U'' shape. It is expected that the ``U''
shape will be the easiest one to detect because complex shape tends to
be concealed by cluster of data points.

@@

Four different distribution were used to generate \(X_{raw}\) as shown
in Figure \ref{fig:different-dist}. The uniform and the normal
distribution are symmetric and commonly assumed in statistical models.
The adjusted log-normal distribution provides skewed density, while the
discrete uniform distribution provides discreteness in residual plot,
which could enrich the pool of visual patterns.

@@

Figure \ref{fig:example-poly-lineup} shows one of the lineups used in
experiment I. This lineup was produced under \(j = 6\) and
\(X_{raw} \sim N(0.0.3^2)\). The actual data plot location was four. All
five subjects correctly identified the actual data plot for this lineup.

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-poly-lineup-1} \caption{Lineup poly-24 in experiment I. Can you spot the most different plot? \label{fig:example-poly-lineup}}\label{fig:example-poly-lineup}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/example-heter-lineup-1} \caption{Lineup heter-169 in experiment II. Can you spot the most different plot? \label{fig:example-heter-lineup}}\label{fig:example-heter-lineup}
\end{figure}

\begin{verbatim}
##  [1]  TRUE FALSE  TRUE  TRUE FALSE FALSE  TRUE  TRUE  TRUE  TRUE  TRUE
\end{verbatim}

An example lineup of this model is shown in Figure
\ref{fig:example-heter-lineup} with \(a = -1\) and
\(X_{raw} \sim U(-1, 1)\). The actual data plot location was 15. 8 out
of 11 subjects correctly identified the actual data plot for this
lineup.

@@

\hypertarget{controlling-the-strength-of-the-signal}{%
\subsubsection{Controlling the strength of the
signal}\label{controlling-the-strength-of-the-signal}}

Other than the parameters summarised in Table
\ref{tab:model-parameter-table}, three additional parameters \(n\),
\(\sigma\) and \(b\) were used to control the strength of the signal in
experiment I and II so that different difficulty levels of lineups were
generated, and therefore, the estimated power curve would be smooth and
continuous. Three different sample sizes were selected (n = 50, 100,
300). In general, with fewer data points shown in a residual plot, the
chance of detecting the visual pattern will decrease. This can be
observed from Figure

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/different-n-1} \caption{Example residual plots of the linear regression model used in experiment I. Three different values of $n$ are included in the experiment to control the strength of the signal.}\label{fig:different-n}
\end{figure}
